DATA 621 - Business Analytics and Data Mining

### 2020-10-28 HW3/Older files/HW3_Final_AA.Rmd
Clean up directory as we are getting close

```{r sumstat}
# DATA EXPLORATION 
summarystat = stby(data = df, INDICES = df$target, FUN = psych::describe)
kable(summarystat[[1]][-13,-c(1,7)], 
      caption = "Descriptive Statistics: Crime Rate > Median", 
      digit = 2L)
kable(summarystat[[2]][-13,-c(1,7)], 
      caption = "Descriptive Statistics: Crime Rate < Median", 
      digit = 2L)
```

## Data Distribution

For each predictors, we computed and drew kernel density estimate to understand 
their distribution. Using density plots, we show how predictors are distributed 
between areas where crime rate is higher than the median, i.e. blue, and areas 
where crime rate is below the median, i.e. red. Our focus is to understand 
variables that highlight large variations between the two groups.

```{r density, fig.width=8}
# density plots
DT[, IDX := .I]
DTM <- melt(DT, id.vars = c('IDX', 'target'), measure.vars = 1:12,
            variable.name = 'metric', value.name = 'value')
ggplot(DTM, aes(x = value, fill = target, color = target)) +
  geom_density(alpha = 0.4, show.legend = FALSE) +
  facet_wrap(~ metric, scales = 'free') 
```

@@

```{r boxplots, fig.width=8}
# boxplots

ggplot(DTM, aes(x = metric, y = value)) + 
  geom_boxplot(color = "blue", fill = "blue", alpha = 0.2, notch = TRUE,
               notchwidth = 0.8, outlier.colour = "red", outlier.fill = "red",
               outlier.size = 3) + 
  stat_summary(fun.y = mean, color = "red", geom = "point", shape = 16,
               size = 2) + coord_flip() +
  labs(title = "Boxplot of Predictor Variables") + scale_y_log10()

@@

```{r correlation, fig.height=5, fig.width=5}
# correlations
corrplot(cor(df), method = 'ellipse', type = 'lower', order = 'hclust')
```

@@

It could be possible we are seeing a reflection of the noticeable influence 
that socio-economic status has on the crime rate of an area.  Graif, Gladfelter 
and Matthews (2016) state crime is generally concentrated in disadvantaged, 
urban neighborhoods in the United States.  An economic segregation suggests 
that affluent neighborhoods may be further away in terms of distance, and as 
a result, disadvantaged areas are more attractive to crime because the probability 
of success is higher even if the targets are not as profitable. We prove this data, 
showing areas with a higher socio-economics correlating to a lower crime rate.

Fields (2004) states industrial areas correlate with higher levels of nitrogen oxide 
concentrations. It may be an indication of why these areas are less dense with residents 
of a higher status. We see the same trend in `ptratio` since a higher ratio means less 
funding for public institutions, which is common in areas of lower status.

([https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4928692/](Source)).
([https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1247398//](Source)).

@@

# DATA PREPARATION

As there are no missing values, we have no need to impute. Instead, different preparation
techniques are utilized for different models.

## Models 1 & 2

The variable `chas` will be removed due to having statistically insignificant correlation 
with almost all the variables in the dataset.  Next, outliers are tempered by capping below 
the 5th percentile and above the 95th percentile.

```{r model1_2_adjust}
new_df <- subset(df, select = -chas)

# Cap all observations at their 5th and 95th percentiles
low5 <- apply(new_df, 2, quantile, prob = 0.05)
up95 <- apply(new_df, 2, quantile, prob = 0.95)
for (i in seq_along(new_df)) {
  new_df[, i] <- pmin(new_df[, i], up95[i])
  new_df[, i] <- pmax(new_df[, i], low5[i])
}
dens_ageO <- ggplot(new_df, aes(age)) + geom_density()
dens_lstatO <- ggplot(new_df, aes(lstat)) + geom_density()
dens_rmO <- ggplot(new_df, aes(rm)) + geom_density()
```

@@

```{r m12BC}
ageBC <- boxcoxfit(new_df$age)
lstatBC <- boxcoxfit(new_df$lstat)
rmBC <- boxcoxfit(new_df$rm)
```

@@

```{r m12applyBC}
new_df$age <- new_df$age ^ ageBC$lambda
new_df$lstat <- new_df$lstat ^ lstatBC$lambda
new_df$rm <- new_df$rm ^ rmBC$lambda
```

@@

```{r m12BC_plot}
dens_age <- ggplot(new_df, aes(age)) + geom_density()
dens_lstat <- ggplot(new_df, aes(lstat)) + geom_density()
dens_rm <- ggplot(new_df, aes(rm)) + geom_density()
grid.arrange(dens_ageO, dens_lstatO, dens_rmO, dens_age, dens_lstat, dens_rm,
             layout_matrix = cbind(c(0, 1, 2), c(3, 4, 5)))
```

## Model 3

For model 3, we chose very little data manipulation, as there may be valuable 
information in the outliers.  A reasonable model may be able to extract 
something from `chas`.  Some variables, such as `chas` and `rad` (an index of 
accessibility) are converted to a factor. Additionally, `rad` probably ordered. 
Given there are now two factors in the predictors, we require creation of dummy 
variables. Lastly, the **labels** of `target` will be changed from integer to 
character for compliance with some of R's naming conventions.  Based on the 
definitions, 0 will be mapped to *below* and 1 to *above*.

# BUILD MODELS

First, we will split the training data into two sets. This will allow us to
perform a cross validation scheme on the models to tune parameters for optimum
performance.

```{r m12_tts}
# BUILD MODELS
# train-test split

crime_df <- new_df %>%
  mutate(target = as.factor(target)) %>% 
  mutate_if(is.integer, as.numeric)

# This is so Model 3 can use the same train/test split as Models 1 and 2
crime_df$id <- seq_len(nrow(crime_df))

# Changing the text above caused changes in output as fewer command were run. It
# is ALWAYS a good idea to set a seed IMMEDIATELY before any sampling in a
# Markdown file to ensure repeatability.

set.seed(7104)
data_split <- initial_split(crime_df, 
                            strata = target, 
                            prop = 0.8)
train_df <- training(data_split)
test_df <- testing(data_split)

# Capture the indices and remove the no-longer-necessary column
tstIDX <- test_df$id
crime_df <- subset(crime_df, select = -id)
train_df <- subset(train_df, select = -id)
test_df <- subset(test_df, select = -id)
```

## Model 1
### Base Model

We will start with a simple logistic regression model to serve as a baseline.
This includes all variables in the dataset.

```{r baseModel1}
baseModel <- glm(target~., family = binomial, data = train_df)
kable(summary(baseModel)$coefficients, digits = 3L,
      caption = 'Base Model Logistic Regression Output')
```

@@

### Enhanced Model

We will use **backwards stepwise regression** to remove features that are not
statistically significant in predicting the target. The result is a model that
includes the following features: `zn`, `nox`, `age`, `dis`, `rad`, `tax`,
`ptratio`, `lstat`, and `medv`.

### Coefficient Discussion

It is important to note that the coefficients from the model are predicting 
whether or not the target variable is a **1** (the crime rate *above* the 
median value). Additionally, since the numeric coefficients are relative to 
the range of values that the variable encompasses it is possible to have a 
coefficient that seems small when we look at the absolute magnitude, but that 
actually has a very strong effect when applied to the data.

```{r model1_coeff}
# 10-fold cross validation
train_control <- trainControl(method = "cv", number = 10)

# train the model on training set
model1 <- train(target ~ zn + nox + age + dis + rad + tax + ptratio + lstat + medv,
               data = train_df,
               trControl = train_control,
               method = "glm",
               family=binomial())
kable(summary(model1)$coefficients, digits = 3L,
      caption = 'Backwards Model Regression Output')
```

@@

The rest of the variables have positive coefficients and many of these are 
expected. For example, in our exploratory data analysis, we noticed that the 
`nox` variable (nitrogen oxides concentration (parts per 10 million)) has the 
greatest positive correlation with the target variable. This is surprising!  
We had not thought that the nitrogen oxide concentration would have as large 
of an impact on the crime rate. According to Colorado State University (2019), 
there may actually be a strong relationship between the two.  The team of 
researchers found strong correlation between exposure to air pollution and 
aggressive behavior.

([https://www.sciencedaily.com/releases/2019/10/191003114007.htm](Source)).

## Model 2
### Logistic Classifier

```{r m2setup}
cv_folds <- vfold_cv(train_df, v = 10, repeats = 1)

@@

crime_wf<- workflow() %>%
  add_recipe(crime_recipe)
```

```{r m2svm}
logit_specs <-
  logistic_reg(
    penalty = tune(),
    mixture = tune()
    ) %>%
  set_engine("glm") %>%
  set_mode("classification")

@@

Tuning the models for best parameters.

```{r m2stack}
logit_wf <-
  crime_wf %>%
  add_model(logit_specs)

crtl_grid<-control_grid(verbose = FALSE)

@@

```{r m2predict}
best_model_auc <- select_best(logit_results, "roc_auc")
final_auc<-
  finalize_workflow(
    logit_wf,
    best_model_auc
)

final_model_pred<-
  final_auc %>%
  last_fit(data_split) %>%
  collect_predictions()

Computing the metrics for the logit model:

```{r m2Stats}
m2_roc<- yardstick::roc_auc(
  final_model_pred,
  truth = target,
  contains(".pred_1")
  )

@@

m2_precise<- yardstick::precision(
  final_model_pred,
  truth = target,
  estimate = .pred_class
)
m2_metrics_df<-
  bind_rows(m2_roc, m2_acc, m2_recall, m2_precise)
```

```{r}
logit_wf %>%
  fit(data = train_df) %>%
  pull_workflow_fit() %>%
  vip(geom = "col", aesthetics = list(fill='red4'))
```

@@

## Model 3
```{r m3trainTemp}
DT_train <- DT[!(IDX %in% tstIDX)]
```

Model 3 will be a binary logistic regression allowing for all variables and 
selected pair-wise interactions pruned using both forward and backward stepwise 
regression based on AICc to select the optimal parameters. The captured indices 
of the test set for models 1 & 2 are used to split the data for model 3 as well. 
To address the possibility of over-fitting, we performed five-fold cross-validation 
on the training set.  With the vanilla logistic regression model in R, `glm` with 
a binomial family and a logit link, there are no other hyper parameters available 
for tuning.  More sophisticated logistic regression implementation have these 
parameters for fine-tuning.

Considering *all* interactions would be foolhardy. Given the need for dummy variables, 
there will 19 predictors and \(_{19}C_2 = 171\) possible pairs, too much. Therefore, 
intentional selection of pairs is required.

Looking at the predictors, there are some, which logically may interact with each other, 
such as:   
* `indus`
	* The concentration of non-retail businesses in the area
* `nox`
    		* Nitrous Oxide is a common industrial pollutant.

@@

```{r m3Clust, fig.height=6}
predClust <- hclust(dist(t(DT_train[, -c('target', 'IDX')]), diag = TRUE))
plot(predClust, xlab = "Raw Predictors")
```

It seems that there may be reason to consider an interaction between `chas` and
`nox`. Can that many industrial plants be on the river?

Similarly with `indus` and `lstat`. Higher-cost homes tend to be further away from the traffic
and pollution of industrial areas.

@@

```{r m3dataPrep}
DT[, `:=`(chas = factor(chas, labels = c('OffRiver', 'OnRiver')),
          rad = factor(rad),
          target = factor(target, labels = c('Below', 'Above')))]
DT_test <- DT[IDX %in% tstIDX]
DT_train <- DT[!(IDX %in% tstIDX)]
```
```{r m3build}
trC <- trainControl(method = 'none', classProbs = TRUE,
                    summaryFunction = twoClassSummary)
dVars <- dummyVars(target ~ indus * nox +
                     indus * lstat +
                     chas * nox +
                     ptratio * medv +
                     rm * dis + . - IDX, data = DT_train, fullRank = TRUE)
DTtrnx <- predict(dVars, newdata = DT_train)
DTtrny <- DT_train$target
set.seed(1)
m3Fit <- train(x = DTtrnx, y = DTtrny, method = 'glmStepAIC', trControl = trC,
               family = 'binomial', trace = 0, direction = 'both')
```

### Coefficient Discussion
```{r m3Sum}
kable(summary(m3Fit$finalModel)$coefficients, digits = 3L,
             caption = 'Stepwise Interaction-Allowed Logistic Regression Output')
```

@@

Lastly, while `rad.1` is the base case, `rad.2` and `rad.6` have disappeared as 
well. This indicates that both `rad = 2` and `rad = 6` should be considered 
indistinguishable from `rad = 1`.

```
                    Df Deviance    AIC
<none>                   90.565 122.56
+ dis                1   89.191 123.19
+ zn                 1   89.558 123.56
+ tax                1   89.700 123.70
- ptratio            1   93.763 123.76
+ age                1   90.156 124.16
+ `nox:chasOnRiver`  1   90.224 124.22
- `rm:dis`           1   94.254 124.25
+ chas.OnRiver       1   90.273 124.27
+ rad.6              1   90.315 124.31
- lstat              1   94.364 124.36
+ `ptratio:medv`     1   90.494 124.49
+ rad.2              1   90.554 124.55
- rm                 1   95.001 125.00
- `indus:lstat`      1   95.633 125.63
- `indus:nox`        1   97.351 127.35
- indus              1   97.597 127.60
- medv               1   98.568 128.57
- rad.3              1  101.594 131.59
- rad.5              1  105.594 135.59
- rad.7              1  118.832 148.83
- nox                1  134.786 164.79
- rad.8              1  149.967 179.97
- rad.4              1  152.757 182.76
- rad.24             1  152.797 182.80
```

@@

```{r comparison}
compTable <- data.frame(Models = c('Model 1', 'Model 2', 'Model 3'),
                        ACC = double(3),
                        F1 = double(3),
                        AUC = double(3))
```

@@

# Metrics
m1_roc <- yardstick::roc_auc(
  model1_results,
  truth = target,
  pred0 # select the prob class that corresponds to first level of target
)

@@

m1CM <- confusionMatrix(model1_preds, model1_results$target,
                        mode = 'prec_recall')
compTable[1, 2] <- m1_acc$.estimate
compTable[1, 3] <- 2 * m1_precise$.estimate * m1_recall$.estimate /
  (m1_precise$.estimate + m1_recall$.estimate)
compTable[1, 4] <- m1_roc$.estimate
m1CM$table
```

@@

```{r m2TestSet}
final_model_pred %>%
  conf_mat(truth = target, estimate = .pred_class)

@@

compTable[2, 2] <- m2.metrics[1, 3]
compTable[2, 3] <- 2 * m2_precise$.estimate * m2_recall$.estimate /
  (m2_precise$.estimate + m2_recall$.estimate)
compTable[2, 4] <- m2.metrics[2, 3]
```

@@

```{r m3TestSet}
tstdVars <- dummyVars(target ~ indus * nox +
                        indus * lstat +
                        chas * nox +
                        ptratio * medv +
                        rm * dis + . - IDX, data = DT_test, fullRank = TRUE)
DTtstx <- predict(tstdVars, newdata = DT_test)
DTtsty <- DT_test$target
m3Pred <- predict(m3Fit, newdata = DTtstx)
m3PredP <- predict(m3Fit, newdata = DTtstx, type = 'prob')
m3CM <- confusionMatrix(m3Pred, DTtsty, mode = 'prec_recall')
m3CM$table
compTable[3, 2] <- sum(diag(m3CM$table)) / sum(m3CM$table)
compTable[3, 3] <- m3CM$byClass[7]
compTable[3, 4] <- yardstick::roc_auc_vec(DTtsty, m3PredP[, 1])
```

### Selected Performance Statistics
```{r selperfstat}
kable(compTable, digits = 3L, caption = 'Model Test Results')
```

@@

```{r evalPred}
setDT(eval)
neval <- dim(eval)[[1]]
eval[, `:=`(chas = factor(chas, labels = c('OffRiver', 'OnRiver')),
          rad = factor(rad))]
evalVars <- dummyVars( ~ indus * nox + indus * lstat + chas * nox + 
                         ptratio * medv + rm * dis + .,
                       data = eval, fullRank = TRUE)
evaltstx <- predict(evalVars, newdata = eval)
evalPred <- predict(m3Fit, newdata = evaltstx)
evalPredP <- predict(m3Fit, newdata = evaltstx, type = 'prob')
predTable <- data.frame(TestID = seq_len(neval),
                        PredictedClass = ifelse(evalPred == "Below", 0, 1),
                        Prob_Is_0 = evalPredP$Below,
                        Prob_Is_1 = evalPredP$Above)
kable(predTable, digits = 3L, caption = "Predicted Classes and Probabilities")
```

# APPENDIX
The code chunks below represent the R code called in order during the analysis.
They are reproduced in the appendix for review and comment.

```{r appendix, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```
```{r loadData}
```
```{r sumstat}
```
```{r density}
```
```{r boxplot}
```
```{r correlation}
```
```{r model1_2_adjust}
```
```{r m12BC}
```
```{r m12applyBC}
```
```{r m12BC_plot}
```
```{r model1_coeff}
```
```{r m2setup}
```
```{r m2svm}
```
```{r m2stack}
```
```{r m2predict}
```
```{r m2Stats}
```
```{r m3trainTemp}
```
```{r m3Clust}
```
```{r m3dataPrep}
```
```{r m3Sum}
```
```{r comparison}
```
```{r model1_perf}
```
```{r m2TestSet}
```
```{r m3TestSet}
```
```{r selperfstat}
```
```{r evalPred}
```

>>>>

```{r sumstat}
# Summary Statistics
summarystat = stby(data = df, INDICES = df$target, FUN = psych::describe)
kable(summarystat[[1]][-13,-c(1,7)], 
      caption = "Descriptive Statistics: Crime Rate > Median", 
      digit = 2L)
kable(summarystat[[2]][-13,-c(1,7)], 
      caption = "Descriptive Statistics: Crime Rate < Median", 
      digit = 2L)
```

## Data Distribution
For each predictors, we computed and drew kernel density estimate to understand 
their distribution. Using density plots, we show how predictors are distributed 
between areas where crime rate is higher than the median, i.e. blue, and areas 
where crime rate is below the median, i.e. red. Our focus is to understand 
variables that highlight large variations between the two groups.

```{r density, fig.width=8}
# Density plots
DT[, IDX := .I]
DTM <- melt(DT, id.vars = c('IDX', 'target'), measure.vars = 1:12,
            variable.name = 'metric', value.name = 'value')
ggplot(DTM, aes(x = value, fill = target, color = target)) +
  geom_density(alpha = 0.4, show.legend = FALSE) +
  facet_wrap(~ metric, scales = 'free') 
```

@@

```{r boxplots, fig.width=8}
# Boxplots 
ggplot(DTM, aes(x = metric, y = value)) + 
  geom_boxplot(color = "blue", fill = "blue", alpha = 0.2, notch = TRUE,
               notchwidth = 0.8, outlier.colour = "red", outlier.fill = "red",
               outlier.size = 3) + 
  stat_summary(fun.y = mean, color = "red", geom = "point", shape = 16,
               size = 2) + coord_flip() +
  labs(title = "Boxplot of Predictor Variables") + scale_y_log10()

@@

```{r correlation, fig.height=5, fig.width=5}
# Corrgram
corrplot(cor(df), method = 'ellipse', type = 'lower', order = 'hclust')
```

@@

It could be possible we are seeing a reflection of the noticeable influence 
that socio-economic status has on the crime rate of an area. Graif, Gladfelter 
and Matthews (2014) state crime is generally concentrated in disadvantaged,
urban neighborhoods in the United States. An economic segregation suggests that
affluent neighborhoods may be further away in terms of distance, and as a
result, disadvantaged areas are more attractive to crime because the probability
of success is higher even if the targets are not as profitable. We prove this
data, showing areas with a higher socio-economics correlating to a lower crime
rate.

Fields (2004) states industrial areas correlate with higher levels of nitrous
oxide concentrations. It may be an indication of why these areas are less dense
with residents of a higher status. We see the same trend in `ptratio` since a
higher ratio means less funding for public institutions, which is common in
areas of lower status.

@@

# DATA PREPARATION
As there are no missing values, we have no need to impute. Further different
preparation techniques are utilized for different models.

## Models 1 & 2
The variable `chas` will be removed due to having statistically insignificant
correlation with almost all the variables in the dataset.  Next, outliers are
tempered by capping below the 5th percentile and above the 95th percentile.

```{r model1_2_adjust}
# Model 1&2 Prep: Remove chas
new_df <- subset(df, select = -chas)

# Model 1&2 Prep:  Cap all observations at their 5th and 95th percentiles
low5 <- apply(new_df, 2, quantile, prob = 0.05)
up95 <- apply(new_df, 2, quantile, prob = 0.95)
for (i in seq_along(new_df)) {
  new_df[, i] <- pmin(new_df[, i], up95[i])
  new_df[, i] <- pmax(new_df[, i], low5[i])
}

# Model 1&2 Prep: Store density plots prior to Box-Cox transform
dens_ageO <- ggplot(new_df, aes(age)) + geom_density()
dens_lstatO <- ggplot(new_df, aes(lstat)) + geom_density()
dens_rmO <- ggplot(new_df, aes(rm)) + geom_density()
```

@@

```{r m12BC}
# Model 1&2 Prep:  Calculate Box-Cox lambdas
ageBC <- boxcoxfit(new_df$age)
lstatBC <- boxcoxfit(new_df$lstat)
rmBC <- boxcoxfit(new_df$rm)
```

@@

```{r m12applyBC}
# Model 1&2 Prep: Apply the Box-Cox transform
new_df$age <- new_df$age ^ ageBC$lambda
new_df$lstat <- new_df$lstat ^ lstatBC$lambda
new_df$rm <- new_df$rm ^ rmBC$lambda
```

@@

```{r m12BC_plot}
# Model 1&2 Prep: Graphically display the effect of the Box-Cox transform
dens_age <- ggplot(new_df, aes(age)) + geom_density()
dens_lstat <- ggplot(new_df, aes(lstat)) + geom_density()
dens_rm <- ggplot(new_df, aes(rm)) + geom_density()
grid.arrange(dens_ageO, dens_lstatO, dens_rmO, dens_age, dens_lstat, dens_rm,
             layout_matrix = cbind(c(0, 1, 2), c(3, 4, 5)))
```

## Model 3
For model 3, we chose very little data manipulation, as there may be valuable 
information in the outliers. A reasonable model may be able to extract something
from `chas`. Some variables, such as `chas` and `rad`, an index of accessibility,
are converted to a factor. Additionally, while `rad` may have an ordering, it
will be converted to a normal factor, as ordered factors are processed via
orthogonal polynoial regression, which is out of the scope of this class. Given
that there are now two factors in the predictors, we require creation of dummy 
variables. Lastly, the **labels** of `target` will be changed from integer to 
character for compliance with some of R's naming conventions.  Based on the 
definitions, 0 will be mapped to *Below* and 1 to *Above*.

# BUILD MODELS
First, we will split the training data into two sets. This will allow us to
perform a cross validation scheme on the models to tune parameters for optimum
performance.

```{r m12_tts}
# Model 1: Split the data into training and testing sets
crime_df <- new_df %>%
  mutate(target = as.factor(target)) %>% 
  mutate_if(is.integer, as.numeric)

# Model 1: This is so Model 3 can use the same train/test split as Models 1 & 2
crime_df$id <- seq_len(nrow(crime_df))

# Changing the text above caused changes in output as fewer command were run. It
# is ALWAYS a good idea to set a seed IMMEDIATELY before any sampling in a
# Markdown file to ensure repeatability.
set.seed(7104)
data_split <- initial_split(crime_df, 
                            strata = target, 
                            prop = 0.8)
train_df <- training(data_split)
test_df <- testing(data_split)

# Model 1: Capture the indices for Model 3 then remove the uneeded column
tstIDX <- test_df$id
crime_df <- subset(crime_df, select = -id)
train_df <- subset(train_df, select = -id)
test_df <- subset(test_df, select = -id)
```

## Model 1
### Base Model
We will start with a simple logistic regression model to serve as a baseline.
This includes all variables in the dataset.

```{r baseModel1}
# Base Model 1
baseModel <- glm(target~., family = binomial, data = train_df)
kable(summary(baseModel)$coefficients, digits = 3L,
      caption = 'Base Model Logistic Regression Output')
```

@@

### Enhanced Model
We will use **backwards stepwise regression** to remove features that are not
statistically significant in predicting the target. The result is a model that
includes the following features: `zn`, `nox`, `age`, `dis`, `rad`, `tax`,
`ptratio`, `lstat`, and `medv`.

### Coefficient Discussion
It is important to note that the coefficients from the model are predicting 
whether or not the target variable is a **1** (the crime rate *above* the 
median value). Additionally, since the numeric coefficients are relative to 
the range of values that the variable encompasses it is possible to have a 
coefficient that seems small when we look at the absolute magnitude, but that 
actually has a very strong effect when applied to the data.

```{r model1_coeff}
# Model 1: 10-fold cross validation
train_control <- trainControl(method = "cv", number = 10)

# Model 1: Train the model on training set
model1 <- train(target ~ zn + nox + age + dis + rad + tax + ptratio + lstat +
                  medv,
                data = train_df,
                trControl = train_control,
                method = "glm",
                family=binomial())
kable(summary(model1)$coefficients, digits = 3L,
      caption = 'Backwards Model Regression Output')
```

@@

The rest of the variables have positive coefficients and many of these are 
expected. For example, in our exploratory data analysis, we noticed that the 
`nox` variable (nitrogen oxides concentration (parts per 10 million)) has the 
greatest positive correlation with the target variable. This is surprising!  
We had not thought that the nitrogen oxide concentration would have as large 
of an impact on the crime rate. According to an article in Science Daily (2019),
researchers at Colorado State University found that there may actually be a
strong correlation between exposure to air pollution and aggressive behavior.

## Model 2
### Logistic Classifier
The first step is to setup a workflow which encapsulates the logistic regression
and cross-validation.

```{r m2setup}
# Model 2: Setup
cv_folds <- vfold_cv(train_df, v = 10, repeats = 1)

@@

crime_wf <- workflow() %>%
  add_recipe(crime_recipe)
```
```{r m2logit}
# Model 2: Call logistic regression
logit_specs <-
  logistic_reg(
    penalty = tune(),
    mixture = tune()
    ) %>%
  set_engine("glm") %>%
  set_mode("classification")

@@

Once the workflow is set up, the model is tuned to return the best parameters.

```{r m2stack}
# Model 2: Setup tuning grid
logit_wf <-
  crime_wf %>%
  add_model(logit_specs)

crtl_grid <-control_grid(verbose = FALSE)

@@

The metric used to select the "best" model will be the area under the receiver
operating curve (AUC).

```{r m2predict}
# Model 2: Select AUC as optimizing metric
best_model_auc <- select_best(logit_results, "roc_auc")
final_auc <-
  finalize_workflow(
    logit_wf,
    best_model_auc
)

final_model_pred <-
  final_auc %>%
  last_fit(data_split) %>%
  collect_predictions()
```

Once the algorithms select a best model, the test metrics can be computed.

```{r m2Stats}
# Model 2: Compute metrics on test set (done here instead of below)
m2_roc<- yardstick::roc_auc(
  final_model_pred,
  truth = target,
  contains(".pred_1")
  )

@@

m2_precise<- yardstick::precision(
  final_model_pred,
  truth = target,
  estimate = .pred_class
)

m2_metrics_df<-
  bind_rows(m2_roc, m2_acc, m2_recall, m2_precise)
```

```{r m2VI}
# Model 2: Plot variable importance
logit_wf %>%
  fit(data = train_df) %>%
  pull_workflow_fit() %>%
  vip(geom = "col", aesthetics = list(fill='red4'))
```

@@

## Model 3
Model 3 will be a binary logistic regression allowing for all variables and 
selected pair-wise interactions pruned using both forward and backward stepwise 
regression based on AICc to select the optimal parameters. The captured indices 
of the test set for models 1 & 2 are used to split the data for model 3 as well. 
With the vanilla logistic regression model in R, `glm` with a binomial family
and a logit link, there are no hyper parameters available for tuning. More
sophisticated logistic regression implementation have these parameters for
fine-tuning. Lastly, the stepAIC method which will be used to select the best
model from an information-criterion perspective, prevents the use of
cross-validation.

Considering *all* interactions would be foolhardy. Given the need for dummy
variables, there will 19 predictors and \(_{19}C_2 = 171\) possible pairs---too
much. Therefore, intentional selection of pairs is required.

Looking at the predictors, there are some, which logically may interact with
each other, such as:

* `indus`
	* The concentration of non-retail businesses in the area
* `nox`
    		* Nitrous Oxide is a common industrial pollutant.

@@

```{r m3Clust, fig.height=6}
# Model 3: Create cluster plot of features
DT_train <- DT[!(IDX %in% tstIDX)]
predClust <- hclust(dist(t(DT_train[, -c('target', 'IDX')]), diag = TRUE))
plot(predClust, xlab = "Raw Predictors")
```

It seems that there may be reason to consider an interaction between `chas` and
`nox`. Can that many industrial plants be on the river? Similarly with `indus`
and `lstat`. Higher-cost homes tend to be further away from the traffic and
pollution of industrial areas.

@@

```{r m3build}
# Model 3: Data Preparation
DT[, `:=`(chas = factor(chas, labels = c('OffRiver', 'OnRiver')),
          rad = factor(rad),
          target = factor(target, labels = c('Below', 'Above')))]
DT_test <- DT[IDX %in% tstIDX]
DT_train <- DT[!(IDX %in% tstIDX)]

# No CV, use precision/recall
trC <- trainControl(method = 'none', classProbs = TRUE,
                    summaryFunction = twoClassSummary)

# Model 3: Create dummy variables
dVars <- dummyVars(target ~ indus * nox +
                     indus * lstat +
                     chas * nox +
                     ptratio * medv +
                     rm * dis + . - IDX, data = DT_train, fullRank = TRUE)
DTtrnx <- predict(dVars, newdata = DT_train)
DTtrny <- DT_train$target

# Model 3: Train Model
set.seed(1)
m3Fit <- train(x = DTtrnx, y = DTtrny, method = 'glmStepAIC', trControl = trC,
               family = 'binomial', trace = 0, direction = 'both')
```

### Coefficient Discussion
```{r m3Sum}
# Model 3: Model summary
kable(summary(m3Fit$finalModel)$coefficients, digits = 3L,
             caption = 'Stepwise Interaction-Allowed Logistic Regression Output')
```

@@

Lastly, while `rad.1` is the base case, `rad.2` and `rad.6` have disappeared as 
well. This indicates that both `rad = 2` and `rad = 6` should be considered 
indistinguishable from `rad = 1`.
```
                    Df Deviance    AIC
<none>                   90.565 122.56
+ dis                1   89.191 123.19
+ zn                 1   89.558 123.56
+ tax                1   89.700 123.70
- ptratio            1   93.763 123.76
+ age                1   90.156 124.16
+ `nox:chasOnRiver`  1   90.224 124.22
- `rm:dis`           1   94.254 124.25
+ chas.OnRiver       1   90.273 124.27
+ rad.6              1   90.315 124.31
- lstat              1   94.364 124.36
+ `ptratio:medv`     1   90.494 124.49
+ rad.2              1   90.554 124.55
- rm                 1   95.001 125.00
- `indus:lstat`      1   95.633 125.63
- `indus:nox`        1   97.351 127.35
- indus              1   97.597 127.60
- medv               1   98.568 128.57
- rad.3              1  101.594 131.59
- rad.5              1  105.594 135.59
- rad.7              1  118.832 148.83
- nox                1  134.786 164.79
- rad.8              1  149.967 179.97
- rad.4              1  152.757 182.76
- rad.24             1  152.797 182.80
```

@@

```{r comparison}
# Create table to hold model comparison statistics
compTable <- data.frame(Models = c('Model 1', 'Model 2', 'Model 3'),
                        ACC = double(3),
                        F1 = double(3),
                        AUC = double(3))
```

@@

# Model 1 Metrics
m1_roc <- yardstick::roc_auc(
  model1_results,
  truth = target,
  pred0 # select the prob class that corresponds to first level of target
)

@@

# Model 1 Confusion Matrix
m1CM <- confusionMatrix(model1_preds, model1_results$target,
                        mode = 'prec_recall')
m1CM$table

# Model 1 Populate comparison table
compTable[1, 2] <- m1_acc$.estimate
compTable[1, 3] <- 2 * m1_precise$.estimate * m1_recall$.estimate /
  (m1_precise$.estimate + m1_recall$.estimate)
compTable[1, 4] <- m1_roc$.estimate
```

@@

```{r m2TestSet}
# Model 2 calculate metrics and print confusion matrix
final_model_pred %>%
  conf_mat(truth = target, estimate = .pred_class)

@@

# Model 2 Populate comparison table
compTable[2, 2] <- m2.metrics[1, 3]
compTable[2, 3] <- 2 * m2_precise$.estimate * m2_recall$.estimate /
  (m2_precise$.estimate + m2_recall$.estimate)
compTable[2, 4] <- m2.metrics[2, 3]
```

@@

```{r m3TestSet}
# Model 3: Process test set
tstdVars <- dummyVars(target ~ indus * nox +
                        indus * lstat +
                        chas * nox +
                        ptratio * medv +
                        rm * dis + . - IDX, data = DT_test, fullRank = TRUE)
DTtstx <- predict(tstdVars, newdata = DT_test)
DTtsty <- DT_test$target

# Model 3: Predict on test set
m3Pred <- predict(m3Fit, newdata = DTtstx)
m3PredP <- predict(m3Fit, newdata = DTtstx, type = 'prob')

# Model 3: Confusion Matrix
m3CM <- confusionMatrix(m3Pred, DTtsty, mode = 'prec_recall')
m3CM$table

# Model 3: Populate comparison table
compTable[3, 2] <- sum(diag(m3CM$table)) / sum(m3CM$table)
compTable[3, 3] <- m3CM$byClass[7]
compTable[3, 4] <- yardstick::roc_auc_vec(DTtsty, m3PredP[, 1])
```

### Selected Performance Statistics
```{r selperfstat}
# Display Model Comparison
kable(compTable, digits = 3L, caption = 'Model Test Results')
```

@@

```{r evalPred}
# Selected Model: Process eval file data as per train and test sets
setDT(eval)
neval <- dim(eval)[[1]]
eval[, `:=`(chas = factor(chas, labels = c('OffRiver', 'OnRiver')),
          rad = factor(rad))]
evalVars <- dummyVars( ~ indus * nox + indus * lstat + chas * nox + 
                         ptratio * medv + rm * dis + .,
                       data = eval, fullRank = TRUE)
evaltstx <- predict(evalVars, newdata = eval)

# Selected Model: Predict classes and probabilities
evalPred <- predict(m3Fit, newdata = evaltstx)
evalPredP <- predict(m3Fit, newdata = evaltstx, type = 'prob')

# Selected Model: Display results
predTable <- data.frame(TestID = seq_len(neval),
                        PredictedClass = ifelse(evalPred == "Below", 0, 1),
                        Prob_Is_0 = evalPredP$Below,
                        Prob_Is_1 = evalPredP$Above)
kable(predTable, digits = 3L, caption = "Predicted Classes and Probabilities")
```

# REFERENCES

 * Fields S. (2004). Global nitrogen: cycling out of control.
 *Environmental health perspectives*, *112*(10), A556–A563.
 https://doi.org/10.1289/ehp.112-a556
 * Graif, C., Gladfelter, A. S., & Matthews, S. A. (2014). Urban Poverty and
 Neighborhood Effects on Crime: Incorporating Spatial and Network Perspectives.
 *Sociology compass*, *8*(9), 1140–1155. https://doi.org/10.1111/soc4.12199
 * Colorado State University. (2019, October 3). Exposure to air pollution
 increases violent crime rates. *ScienceDaily*. Retrieved October 26, 2020 from
 www.sciencedaily.com/releases/2019/10/191003114007.htm

# CODE APPENDIX
The code chunks below represent the R code called in order during the analysis.
They are reproduced in the appendix for review and comment.
